### 2019-09-28 paper-two/paper-two.tex
dmn done, sn next

The DMN is comprised of specific anatomy including portions of the mid-line of the
brain, an evolutionarily primitive area related to memory and emotion, and
structures in the cortex, an evolutionarily recent part of the brain containting the
executive and higher order functions. \cite{defaultnetworkanatomy}

These regions are functionally connected and are involved in the neurological basis
of the self, considering the mental states of others, remembering the past, and
imagining the future.

a) The neurological basis of the self —
The regions involved in this are the Posterior Cingulate Cortex (PCC), venteromedial Prefrontal Cortex (vmPFC) and the Inferiour Parietal Lobule (IPL) (1).
Cerebral blood flow and metabolic rate is 40\% higher in the PCC compared to the average across the brain making it one of the most active regions of the brain. (Leech R, Sharp DJ (July 2013). “The role of the posterior cingulate cortex in cognition and disease”. Brain. 137 (Pt 1): 12–32.)
The PCC has been thought to mediate interactions between emotion and memory. It consistently lights up when a person recollects something from their life and depending on the emotional importance of the memory the strength of the lighting up changes. It is also activated by emotional stimuli both positive and negative and acts as an interface between the world and us by gauging the importance of different stimuli(2).

Magnetic resonance Image of the areas of the brain in the Default Mode Network. http://www.frontiersin.org/Neurotrauma/10.3389/fneur.2013.00016/full
The mPFC is divided into venteromedial PreFrontal Cortex (vmPFC) and dorsomedial Prefrontal Cortex (dmPFC). The vmPFC is involved in creating a conceptual self. It is involved in self-related processing and assigns personal significance to self-related information (3).
It is hypothesized that the activation of these three regions of the brain are responsible for providing a sense of self — as a subjective agent in space and time. It functions as a network in which phenomena such as self representations are accessible via the PCC and are gated into conscious awareness by activity in the MPFC which is influenced by changing internal and external demands. The PCC which acts as a brain-wide connectivity hub through which a unitary notion of self is created by a large scale integration of DMN activity (4).
b) Thinking about others —
The Dorsal Medial Prefrontal Cortex (dmPFC) is involved in Theory of the mind(considering the mental states of others), empathy, moral reasoning and altruistic behavior (1,5,6)
c) Autobiographical memory and future simulations —
The DMN is also involved in remembering the past and imagining the future and story comprehension(1).
Salience Network (SN):
Salience: The perceptual quality by which an observable thing stands out relative to its environment
The SN is an intrinsically connected large-scale network anchored in the Anterior Insula (AI) and dorsal Anterior Cingulate Cortex. Both these regions have reached a high degree of specialization in the great apes. It is the collection of the regions in the brain that help decide which stimuli deserves our attention. It acts as a switch between the internally directed DMN and the externally directed Central Executive Network (involved in high level cognitive functions) (7).

@@

Maddock, Richard J.; Garrett, Amy S.; Buonocore, Michael H. (January 2003). “Posterior cingulate cortex activation by emotional words: fMRI evidence from a valence decision task”. Human Brain Mapping. 18 (1): 30–41.
Andrews-Hanna, Jessica R.; Smallwood, Jonathan; Spreng, R. Nathan (2014–05–01). “The default network and self-generated thought: component processes, dynamic control, and clinical relevance”. Annals of the New York Academy of Sciences. 1316(1): 29–52. doi:10.1111/nyas.12360. ISSN 1749–6632. PMC 4039623. PMID 24502540.
Davey CG, Pujol J, Harrison BJ. Mapping the self in the brain’s default mode network. Neuroimage. 2016 May 15;132:390–397.
Isoda, M., \& Noritake, A. (2015). What makes the dorsomedial frontal cortex active during reading the mental states of others?. Neural basis of social learning, social deciding, and other-regarding preferences, 51.
Waytz, A., Zaki, J., \& Mitchell, J. P. (2012). Response of dorsomedial prefrontal cortex predicts altruistic behavior. The Journal of Neuroscience, 32(22), 7646–7650.

@@

\bibitem{thenewsroom}
  Aaron Sorkin.
  \url{https://www.youtube.com/watch?v=2C6h-Yyx9Yk}
  \textit{The Newsroom}.
  HBO, 2012.

>>>>

The DMN is comprised of specific anatomy including portions of the mid-line of the
brain, an evolutionarily primitive area related to memory and emotion, and
structures in the cortex, an evolutionarily recent part of the brain containting the
executive and higher order functions. \cite{defaultnetworkanatomy} These functionally
connected regions are involved in the neurological basis
of the self, considering the mental states of others, remembering the past, and
imagining the future.

\subsubsection{The neurological basis of the self}

The regions involved in an individual's conceptual ``self'' are the Posterior
Cingulate Cortex (PCC), Ventromedial Prefrontal Cortex (vmPFC), and the Inferior
Parietal Lobule (IPL). \cite{defaultnetworkadaptive}

Cerebral blood flow and metabolic rate is 40\%
higher in the PCC compared to the average across the brain, making it one of the most
active regions of the brain. \cite{pccrole}

The PCC has been thought to mediate interactions between emotion and memory. Under
fMRI, it consistently lights up when a person recollects something from their
life. The strength of PCC brain activity in such an experiment varies depending on
the emotional importance of the memory recalled. It is also activated by emotional
stimuli, both positive and negative, and acts as an interface between the external
world and the individual by gauging the importance of different
stimuli. \cite{pccemotion}

[ TODO: MRI of DMN image ] \cite{dmnmri}

The vmPFC is involved in creating a conceptual self by self-related processing
and the assignment of personal significance to self-related
information. \cite{dmnself} The vmPFC is paired with the Dorsomedial Prefrontal
Cortex (dmPFC), which assists in the creation of the conceptual self through an
individual's consideration of ``others'', to construct the Medial Prefrontal Cortex
(mPFC).

[ TODO: explain IPL? ]

It is hypothesized that the activation of these three regions of the brain (PCC,
vmPFC, and IPL) are responsible for providing a sense of self --- a subjective agent
in space and time. It functions as a network in which phenomena such as self
representations are accessible via the PCC [ TODO: ref ] and are gated into conscious awareness by
activity in the mPFC, as influenced by changing internal and external demands. The
PCC acts as a brain-wide connectivity hub, through which a unitary notion of self is
created by a large scale integration of DMN activity. \cite{mappingself}

\subsubsection{Thinking about others}

The \textit{Theory of Mind} is a body of work in philosophy and psychology which
describes an individual's ability to consider the mental states of others, such as
appreciating another individual's false beliefs (knowledge based on incorrect or
outdated information). The Theory of Mind captures not only one's ability to
attribute beliefs, desires, and emotions to others but also to oneself --- and to
differentiate the two. \cite{autistictheoryofmind}

In addition to The Theory of Mind, the dmPFC is involved in empathy, moral reasoning,
and altruistic behavior. \cite{defaultnetworkadaptive,dmpfcothers,dmpfcaltruism}

\subsubsection{Autobiographical memory and future simulations}

The DMN is also involved in remembering the past, imagining the future, and story
comprehension. \cite{defaultnetworkadaptive}

[ TODO: expand? how? ]

\subsection{Salience Network}

\begin{quote}
  \textbf{Salience: The perceptual quality by which an observable thing stands out
    relative to its environment}
\end{quote}

The SN is an intrinsically connected large-scale network anchored in the Anterior Insula (AI) and dorsal Anterior Cingulate Cortex. Both these regions have reached a high degree of specialization in the great apes. It is the collection of the regions in the brain that help decide which stimuli deserves our attention. It acts as a switch between the internally directed DMN and the externally directed Central Executive Network (involved in high level cognitive functions) (7).

@@

@@

\bibitem{dmnmri}
  \textit{Magnetic Resonance Image of the areas of the brain in the Default Mode Network.}
  http://www.frontiersin.org/Neurotrauma/10.3389/fneur.2013.00016/full

\bibitem{pccrole}
  Leech R, Sharp DJ
  \textit{The role of the posterior cingulate cortex in cognition and disease.}
  Brain. 137 (Pt 1): 12–32., July 2013.

\bibitem{pccemotion}
  Maddock, Richard J.; Garrett, Amy S.; Buonocore, Michael H.
  \textit{Posterior cingulate cortex activation by emotional words: fMRI evidence from a
    valence decision task.}
  Human Brain Mapping. 18 (1): 30–41. January 2003.

\bibitem{dmnself}
  Andrews-Hanna, Jessica R.; Smallwood, Jonathan; Spreng, R. Nathan.
  \textit{The default network and self-generated thought: component processes,
    dynamic control, and clinical relevance.}
  Annals of the New York Academy of Sciences. 1316(1):
  29–52. doi:10.1111/nyas.12360. ISSN 1749–6632. PMC 4039623. PMID 24502540. 2014-05-01.

\bibitem{mappingself}
  Davey CG, Pujol J, Harrison BJ.
  \textit{Mapping the self in the brain’s default mode network.}
  Neuroimage. 132:390–397., 2016-05-15

\bibitem{autistictheoryofmind}
  Baron-Cohen, Simon; Leslie, Alan M.; Frith, Uta.
  \textit{Does the autistic child have a ``theory of mind''?.}
  Cognition. 21 (1): 37–46. doi:10.1016/0010-0277(85)90022-8. PMID
  2934210. Pdf. October 1985.

\bibitem{dmpfcothers}
  Isoda, M., \& Noritake, A.
  \textit{What makes the dorsomedial frontal cortex active
    during reading the mental states of others?.}
  Neural basis of social learning, social deciding, and other-regarding preferences,
  51. 2015.

\bibitem{dmpfcaltruism}
  Waytz, A., Zaki, J., \& Mitchell, J. P.
  \textit{Response of dorsomedial prefrontal cortex predicts altruistic behavior.}
  The Journal of Neuroscience, 32(22), 7646–7650. 2012.
